Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2017-07-18 Response.tex

        % draw terminal state on reachability set
        \node[tdplot_rotated_coords,label=above:\(\vecbf{x}_f\)] (xf) at ($ (x0)+(-45:1.5) $) {};
        \filldraw (xf) circle (3pt);

        \node[tdplot_rotated_coords,label=below:\(J\)] at (xf) {};

        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
        % place
    \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Reachability set on a \Poincare section: Pictoral representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
        }
\end{figure}
\end{correction}
\item
    \begin{itshape}
Consider plotting the intersections of manifolds with the Poincare section as continuous curves.  
\end{itshape}

Thank you for the comment.
The manifold intersections with the \Poincare section are plotted as discrete points as the curves were generated in a discrete manner. 
A series of trajectories were generated by perturbing the trajectory at various points along the periodic orbit. 
As a result, we show that the intersection is a discrete approximation of the true intersection of the manifold with the section by using the discrete points rather than a continuous curve.

### 2017-07-18 Response.tex

        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
        % place
    \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Reachability set on a \Poincare section: Pictoral representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
        }
\end{figure}
\end{correction}
\item
    \begin{itshape}
Consider plotting the intersections of manifolds with the Poincare section as continuous curves.  
\end{itshape}

Thank you for the comment.
The manifold intersections with the \Poincare section are plotted as discrete points as the curves were generated in a discrete manner. 
A series of trajectories were generated by perturbing the trajectory at various points along the periodic orbit. 
As a result, we show that the intersection is a discrete approximation of the true intersection of the manifold with the section by using the discrete points rather than a continuous curve.

\item
    \begin{itshape}
Also consider zooming/scaling as necessary to show the details you discuss.
\end{itshape}

Thank you for the comment, however we feel that the figures adequately show the desired information.
The figures are centered around the various intersections of the reachable set and the \Poincare section.
Furthermore, in the case of the much larger geostationary transfer requires it becomes more difficult to zoom in on the fine details.
However, a detailed view is presented which shows the various iterations of the reachability set computations.

### 2017-07-18 Response.tex

Thank you for the comment, however we feel that the figures adequately show the desired information.
The figures are centered around the various intersections of the reachable set and the \Poincare section.
Furthermore, in the case of the much larger geostationary transfer requires it becomes more difficult to zoom in on the fine details.
However, a detailed view is presented which shows the various iterations of the reachability set computations.

\item
    \begin{itshape}
*** Method/Examples
I urge a revisiting of the claims with regards to optimality.  The optimization method is being initialized with *something*, and stationary points are *local* extrema.  More precise claims and assurances about solution quality (and therefore the validity of the computed reachable set boundary) are needed.  
\end{itshape}

\item
    \begin{itshape}
How damaging is the assumption of a fixed travel time?  And how about the specification of a specific departure point on the initial orbit?
\end{itshape}

The fixed travel time is primarily used at this stage to reduce the complexity of the optimal control formulation. 
Furthermore, the reachability set is defined in terms of a fixed terminal time.
A variable terminal travel time would add an additional dimension, time, to the reachable space and add additional complexity to the reachability set visualization.
The fixed travel time does not adversely affect the optimality of the computed trajectories on the reachable set, as the reachable set is defined in terms of a fixed travel time.

The choice of starting from a fixed point on the initial periodic orbit is again mostly used for convience. 
In this case of the periodic orbit the departing trajectories are all along the unstable invariant manifold associated with the periodic orbit. 
The uncontrolled trajecotries associated with the invariant manifolds serve as the ``center'' of each reachability set computation.
The low-thrust propulsion system is then used to enlarge the distance between the uncontrolled and controlled states. 
Starting from a different location on the periodic orbit will not create a qualitatively different situation as the entire invariant manifold also intersects the \Poincare section.
However, different locations along the periodic orbit may have vastly different travel times before intersecting the \Poincare section.
This will require modification of the optimal control problem, with a larger terminal time, but will result in a similar behavior regardless of the location on the intial periodic orbit.

### 2017-07-18 Response.tex

The choice of starting from a fixed point on the initial periodic orbit is again mostly used for convience. 
In this case of the periodic orbit the departing trajectories are all along the unstable invariant manifold associated with the periodic orbit. 
The uncontrolled trajecotries associated with the invariant manifolds serve as the ``center'' of each reachability set computation.
The low-thrust propulsion system is then used to enlarge the distance between the uncontrolled and controlled states. 
Starting from a different location on the periodic orbit will not create a qualitatively different situation as the entire invariant manifold also intersects the \Poincare section.
However, different locations along the periodic orbit may have vastly different travel times before intersecting the \Poincare section.
This will require modification of the optimal control problem, with a larger terminal time, but will result in a similar behavior regardless of the location on the intial periodic orbit.

\item
    \begin{itshape}
In Fig 10b, it is hard to see, but appears that the low-thrust-reachable set is actually *centered* on part of the target set.  Is this related to the existence of the uncontrolled transfer opportunity?  There must be some more clear way to show how the low thrust set expands upon the zero thrust set.  Perhaps beginning with the time horizon of the zero-thrust opportunity, and successively lowering it, would help.  
\end{itshape}

The reachability set is closely spaced in terms of both the position space as well as on the \Poincare section.
This is due to the fact that there is a very low-thrust control input on the spacecraft.
As a result, the trajectories associated with the reachability set are all closely spaced, but more widely seperated from the terminal state without any control input.
In the figure, there is a fixed point labeled \( x_n \) which is the terminal state without any control input, which will also lie on the periodic orbit.
The reachable set, in red, is widely separated from the uncontrolled state due the effect of the low thrust proplusion, however the subsequent spread of the entire reachability set is much more constrained.
The low thrust expands the reachability set from a fixed point to the red elliptical region. 
This is also explained in the manuscript with a similar description of the figure.

### 2017-07-18 Response.tex

\item
    \begin{itshape}
In Fig 10b, it is hard to see, but appears that the low-thrust-reachable set is actually *centered* on part of the target set.  Is this related to the existence of the uncontrolled transfer opportunity?  There must be some more clear way to show how the low thrust set expands upon the zero thrust set.  Perhaps beginning with the time horizon of the zero-thrust opportunity, and successively lowering it, would help.  
\end{itshape}

The reachability set is closely spaced in terms of both the position space as well as on the \Poincare section.
This is due to the fact that there is a very low-thrust control input on the spacecraft.
As a result, the trajectories associated with the reachability set are all closely spaced, but more widely seperated from the terminal state without any control input.
In the figure, there is a fixed point labeled \( x_n \) which is the terminal state without any control input, which will also lie on the periodic orbit.
The reachable set, in red, is widely separated from the uncontrolled state due the effect of the low thrust proplusion, however the subsequent spread of the entire reachability set is much more constrained.
The low thrust expands the reachability set from a fixed point to the red elliptical region. 
This is also explained in the manuscript with a similar description of the figure.

\item
    \begin{itshape}
Perhaps there is also a better way to show what the control is doing, e.g., in some alternative parametrization?  This could help provide assurances that the solution is doing something meaningful and is not just stuck in a weird local minimum.  
\end{itshape}

Thank you for the comment.
The control plots show the planar components of the control thrust. 
Furthermore, the scale has been modified to show the thrust in terms of Newtons for an assumed \SI{500}{\kilo\gram} satellite.
The control plots are meant to demonstrate the control input required by the trajectories which lie on the maximum reachability set. 
As a result, the control input is determined through optimization to maximize the reachability set in a particular direction on the \Poincare section.
This is in contrast to typical optimal control applications where the control input is used directly in the trajectory design process.
Our method seeks to approximate the reachability set, and from this set trajectories are chosen in a systematic method which drives the spacecraft towards a target.

### 2017-07-18 Response.tex

\item
    \begin{itshape}
Perhaps there is also a better way to show what the control is doing, e.g., in some alternative parametrization?  This could help provide assurances that the solution is doing something meaningful and is not just stuck in a weird local minimum.  
\end{itshape}

Thank you for the comment.
The control plots show the planar components of the control thrust. 
Furthermore, the scale has been modified to show the thrust in terms of Newtons for an assumed \SI{500}{\kilo\gram} satellite.
The control plots are meant to demonstrate the control input required by the trajectories which lie on the maximum reachability set. 
As a result, the control input is determined through optimization to maximize the reachability set in a particular direction on the \Poincare section.
This is in contrast to typical optimal control applications where the control input is used directly in the trajectory design process.
Our method seeks to approximate the reachability set, and from this set trajectories are chosen in a systematic method which drives the spacecraft towards a target.

\item
    \begin{itshape}
In the second example, what do you lose by chaining together multiple reachability-based steps?  Aren't you imposing that the trajectory routinely returns to the energy level of the Poincare section? How feasible would it be to conduct a single reachability analysis covering multiple revolutions? And can you use knowledge about directionality of movement on the Poincare section to limit the scope of your reachability search?
\end{itshape}

While the trajectories do return to the energy level of the \Poincare section, there is a transfer of ``energy'' between the states of the vehicle.
It is in this manner that the reachability set is enlarged to eventually include the target state.

Conducting a single reachability analysis becomes more complicated as the larger time of flight increases the sensitivity of the terminal states to variations in the initial costates. 
Furthermore, the larger time of flight creates a much larger and more chaotic reachable set for the analysis.
Through this work it was found that shorter trajectories tended to have more stable and consistent solutions.

### 2017-07-19 Response.tex

    \tdplotdrawarc[orbit hidden]{(center)}{2}{90}{190}{}{};
    \tdplotdrawarc[orbit,<-]{(center)}{2}{-170}{90}{}{};

    \tdplotdrawarc[orbit hidden]{(center)}{3}{90}{199}{}{};
    \tdplotdrawarc[orbit,<-]{(center)}{3}{-161}{90}{}{};

        \end{tikzpicture}
    \end{scaletikzpicturetowidth}
    \caption{Diagram of the \Poincare map: Periodic orbits will appear as fixed points on the \Poincare section \( \Sigma \). Stabiilty of periodic orbits is clearly visible on the section as successive intersections approach or depart a fixed point.}
\end{figure}

\begin{figure}[H]
        \centering 
        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparision: A stable trajectory is used to test the variational and conventional integrators. 
        The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.}
        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{trajectory}}~
        \subcaptionbox{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}: A range of fixed step sizes are used for the variational integrator to approximately match the computational time of \texttt{ODE45}}
        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
\end{figure}

\begin{figure}[H]
        \centering
%       \includegraphics[width=0.5\textwidth]{reachability_set_hand}
        \begin{scaletikzpicturetowidth}{0.5\textwidth}
        \begin{tikzpicture}[tdplot_main_coords,
          poincare/.style={opacity=.2,very thick,fill=blue},
          orbit/.style={very thick,black},
          orbit hidden/.style={very thick,dashed},
          grid/.style={very thin,black},
          axis/.style={->,blue,thick},
          reachability/.style={thick,blue},scale=\tikzscale]

        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
        % place
    \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Reachability set on a \Poincare section: Pictoral representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
            The terminal state, \( x_n\), is the intersection without any control input. 
            Adding a control input allows for the terminal state, \( x_f \), to be displaced by some distance/cost \( J \) as measured on the section.
            We parameterize a specific direction on the section with the angles \( \theta_d\) and seek to maximize the distance between \( x_f \) and \( x_n \).
            Computation of the maximum distance, or reacahbility, for a variety angles gives a discrete approximation of the reachability set.}
\end{figure}

\begin{figure}[H]
        \centering
        \begin{scaletikzpicturetowidth}{1\textwidth}
                \begin{tikzpicture}[scale=\tikzscale]
                    % \draw[help lines] (0,0) grid (20,5) ; %grid

                    \node[label=west:\(\vecbf{x}\)] (x0m) at (0,5) {};
                    \node[] (x1m) at (5,5) {};
                    \node[] (x2m) at (10,5) {};
                    \node[] (x3m) at (15,5) {};
                    \node[] (xnm) at (20,5) {};

                    \node[label=below:\(k_0\)] (t0) at (0,0) {};
                    \node[label=below:\(k_1\)] (t1) at (5,0) {};
                    \node[label=below:\(k_2\)] (t2) at (10,0) {};
                    \node[label=below:\(k_{n-1}\)] (t3) at (15,0) {};
                    \node[label=below:\(k_n\)] (tn) at (20,0) {};

                    % nodes for each subsegment
                    \node[label=left:\(\vecbf{x}_0\)] (x0) at (0,1) {\pgfuseplotmark{*}};
                    \node[label=below left:\(\vecbf{\lambda}_0\)] at (x0) {};

                    \node[label=right:\(\vecbf{x}_1^{-}\)] (x1minus) at (5,2) {\pgfuseplotmark{*}};
                    \node[label=below right:\(\vecbf{\lambda}_1^{-}\)] at (x1minus) {};

                    \node[label=above left:\(\vecbf{x}_1^{+}\)] (x1plus) at (5,3) {\pgfuseplotmark{o}};
                    \node[label=left:\(\vecbf{\lambda}_1^{+}\)] at (x1plus) {};

                    \node[label=right:\(\vecbf{x}_2^{-}\)] (x2minus) at (10,2) {\pgfuseplotmark{*}};
                    \node[label=below right:\(\vecbf{\lambda}_2^{-}\)] at (x2minus) {};

                    \node[label=above left:\(\vecbf{x}_2^{+}\)] (x2plus) at (10,3) {\pgfuseplotmark{o}};
                    \node[label=left:\(\vecbf{\lambda}_2^{+}\)] at (x2plus) {};

                    \node[label=right:\(\vecbf{x}_{n-1}^{-}\)] (x3minus) at (15,2) {\pgfuseplotmark{*}};
                    \node[label=below right:\(\vecbf{\lambda}_{n-1}^{-}\)] at (x3minus) {};

                    \node[label=above left:\(\vecbf{x}_{n-1}^{+}\)] (x3plus) at (15,3) {\pgfuseplotmark{o}};
                    \node[label=left:\(\vecbf{\lambda}_{n-1}^{+}\)] at (x3plus) {};

                    \node[label=right:\(\vecbf{x}_n\)] (xnminus) at (20,3) {\pgfuseplotmark{*}};
                    \node[label=below right:\(\vecbf{\lambda}_n\)] at (xnminus) {};
                    % draw axes
                    \draw [<->,thick] (tn.center) -- (t0.center) -- (x0m.center);

                    % draw segement dividers
                    \draw [dashed,thick] (t1.center) -- (x1m.center);
                    \draw [dashed,thick] (t2.center) -- (x2m.center);
                    \draw [dashed,thick] (t3.center) -- (x3m.center);
                    \draw [dashed,thick] (tn.center) -- (xnm.center);

                    % draw markers at each subsegment

                    % draw the subsegments
                    \draw [->] (x0.center) to [bend left=30] (x1minus.center);
                    \draw [->] (x1plus.center) to [bend left=30] (x2minus.center);
                    % dotted lines here
                    \draw[decorate sep={1mm}{5mm},fill] (x2plus.east) to [bend left=10] (x3minus.west);

                    \draw [->] (x3plus.center) to [bend left=30] (xnminus.center);
                \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Schematic diagram of the multiple shooting method:
        The complete trajectory is split into a nubmer of sub-segments, and additional interior constraints are included to ensure state and costate continunity.
    Splitting the optimal trajectory into short segments reduces the sensitivity of terminal states to variations of the initial states.}
\end{figure}

\begin{figure}[H]
   \centering
   \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstraiting the initial and target orbit.
   Without the low-thrust proplusion, the spacecraft is constrained to the initial periodic orbit. 
   We determine the reachability set to find a transfer trajectory to the target orbit about the moon.}
\end{figure}

\begin{figure}[H]
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{manifold_trajectory} 
                \caption{Position space view of invariant manifold} \label{fig:manifold_trajectory} 
        \end{subfigure}~ %
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{manifold_poincare} 
                \caption{\Poincare intersection of invariant manifold (green), target orbit (blue) and the initial orbit (black)} \label{fig:manifold_poincare} 
        \end{subfigure} 
        \caption{Invariant manifold transfer: An example transfer using the invariant manifolds is shown in both the position and \Poincare spaces.
        The control free transfer from the initial periodic orbit to the target orbit result in a long time of flight. 
    In addition, the manifold only intersects the target orbit on the descending or far side of the moon.}
\end{figure}
\begin{figure}[H]
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{reach_trajectory} 
                \caption{Position space view of reachability set}  
        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{poincare_compare} 
                \caption{\Poincare section view of reachability set}  
        \end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)

        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{reach_transfer} 
                \caption{Transfer trajectory selected from reachabiltiy set viewed in the position space}  
        \end{subfigure}~ 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{control_input_l1} 
                \caption{Control input for the selected transfer trajectory}  
        \end{subfigure}~
        \caption{\( L_1 \) Reachability set transfer: The low thrust propulsion is used to approximate the reachability set starting from the initial periodic orbit over a fixed time horizon.
        The reachability set is shown in the upper two figures in both the position and \Poincare space.
    From this reachabiilty set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
The optimal control to achieve this transfer is shown in the lower right figure.} 
\end{figure}
\end{correction}
\item
    \begin{itshape}
Figure 4 does not show what I think you intend it to show; it looks like two periodic circular orbits.
\end{itshape}

### 2017-07-19 manuscript.tex
Update more captions for the figures
Reviewer 2 Comment 7

        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
        % place
    \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Reachability set on a \Poincare section: Pictoral representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
        \label{fig:reachability_set}}
\end{figure}

@@

                    \draw [->] (x3plus.center) to [bend left=30] (xnminus.center);
                \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Multiple Shooting method\label{fig:multiple shooting}}
\end{figure}

@@

\subsection{Periodic Orbit transfer}\label{sec:periodic_orbit_transfer}
% define problem setup (initial and target orbits)
The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix}^T \) with \( \mu = 0.0125 \).
The target set is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
   \caption{Target Orbit Region}
   \label{fig:moon_orbit}
\end{figure}
\Cref{fig:moon_orbit} shows that the target set remains in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
This type of orbit would be useful for a variety of mission scenarios.
For example, a series of communication satellites could be placed in this type of orbit. 
The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.
The initial set is a planar periodic orbit about \( L_1\), which is generated using the process of differential correction of a linear approximation~\cite{koon2011}.

% invariant manifold transfer
As a source for comparison, the method of using invariant manifolds, introduced in~\cite{koon2011}, is implemented.
As described in~\cref{sec:invariant_manifold}, these invariant manifolds are the set of trajectories that either asymptotically arrive or depart the periodic orbit. 
We generate the unstable manifold associated with the initial planar periodic orbit.
We numerically propagate the unstable manifold forward in time until the trajectories intersect the \Poincare section \( y = 0 \).
\Cref{fig:manifold_trajectory} shows the unstable invariant manifold generated from the initial \( L_1\) periodic orbit. 
The blue points in~\cref{fig:manifold_poincare} are the intersections of the target Moon orbit and the \Poincare section.
The two circular regions are the ascending (right) and descending (left) intersections of the target orbit and \Poincare section.
The green points in~\cref{fig:manifold_poincare} are intersections of the unstable manifold from~\cref{fig:manifold_trajectory} with the \Poincare section \( y = 0 \).
Only a single branch of the invariant manifold intersects with the ascending region of the target orbit.
There are no intersections of the invariant manifold with the descending region of the target orbit.
The numerical values associated with the green points denote the required time of flight along the invariant manifold in non-dimensional units.
The required travel time for a transfer using the unstable invariant manifold is  approximately \( t_f \approx 3.1\) non-dimensional time units.
\begin{figure} 
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{manifold_trajectory} 
                \caption{Invariant Manifold} \label{fig:manifold_trajectory} 
        \end{subfigure}~ %
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{manifold_poincare} 
                \caption{Poincar\`e Section} \label{fig:manifold_poincare} 
        \end{subfigure} 
        \caption{Invariant manifold transfer}
        \label{fig:invariant_manifold_transfer} 
\end{figure}

@@

% discuss reachability set computation
The optimal trajectories, under the influence of the control input \( \vecbf{u} \), are plotted in red in~\cref{fig:reach_trajectory}.
Initially, the spacecraft is assumed to lie on the periodic orbit.
As a result, the intersection of this periodic orbit with the \Poincare section are two points corresponding to the two crossing of the orbit.
We show the control-free intersection, \( \vecbf{x}_n \), of the periodic orbit on the \Poincare section in~\cref{fig:manifold_poincare,fig:poincare_compare}
The use of the continuous low thrust propulsion expands the reachable set to region bounded by the red markers in~\cref{fig:poincare_compare}.
The reachable set is an ellipsoidal region with a major axis aligned along \( \theta \approx \ang{70} \) as compared to a fixed point without any control input.
\begin{figure} 
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{reach_trajectory} 
                \caption{Reachability set trajectories} \label{fig:reach_trajectory} 
        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{poincare_compare} 
                \caption{\Poincare section} \label{fig:poincare_compare} 
        \end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)

        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{reach_transfer} 
                \caption{Final transfer trajectory} \label{fig:reach_transfer} 
        \end{subfigure}~ 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{control_input_l1} 
                \caption{Control input} \label{fig:control_l1} 
        \end{subfigure}~
        \caption{Reachability set transfer}
        \label{fig:reachability_set_transfer} 
\end{figure}

>>>>

        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
        % place
    \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Reachability set on a \Poincare section: Pictoral representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
            The terminal state, \( x_n\), is the intersection without any control input. 
            Adding a control input allows for the terminal state, \( x_f \), to be displaced by some distance/cost \( J \) as measured on the section.
            We parameterize a specific direction on the section with the angles \( \theta_d\) and seek to maximize the distance between \( x_f \) and \( x_n \).
            Computation of the maximum distance, or reacahbility, for a variety angles gives a discrete approximation of the reachability set.\label{fig:reachability_set}}
\end{figure}

@@

                    \draw [->] (x3plus.center) to [bend left=30] (xnminus.center);
                \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Schematic diagram of the multiple shooting method:
        The complete trajectory is split into a nubmer of sub-segments, and additional interior constraints are included to ensure state and costate continunity.
    Splitting the optimal trajectory into short segments reduces the sensitivity of terminal states to variations of the initial states.\label{fig:multiple shooting}}
\end{figure}

@@

\subsection{Periodic Orbit transfer}\label{sec:periodic_orbit_transfer}
% define problem setup (initial and target orbits)
The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix}^T \) with \( \mu = 0.0125 \).
The target set is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstraiting the initial and target orbit.
   Without the low-thrust proplusion, the spacecraft is constrained to the initial periodic orbit. 
   We determine the reachability set to find a transfer trajectory to the target orbit about the moon.}
   \label{fig:moon_orbit}
\end{figure}
\Cref{fig:moon_orbit} shows that the target set remains in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
This type of orbit would be useful for a variety of mission scenarios.
For example, a series of communication satellites could be placed in this type of orbit. 
The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.
The initial set is a planar periodic orbit about \( L_1\), which is generated using the process of differential correction of a linear approximation~\cite{koon2011}.

% invariant manifold transfer
As a source for comparison, the method of using invariant manifolds, introduced in~\cite{koon2011}, is implemented.
As described in~\cref{sec:invariant_manifold}, these invariant manifolds are the set of trajectories that either asymptotically arrive or depart the periodic orbit. 
We generate the unstable manifold associated with the initial planar periodic orbit.
We numerically propagate the unstable manifold forward in time until the trajectories intersect the \Poincare section \( y = 0 \).
\Cref{fig:manifold_trajectory} shows the unstable invariant manifold generated from the initial \( L_1\) periodic orbit. 
The blue points in~\cref{fig:manifold_poincare} are the intersections of the target Moon orbit and the \Poincare section.
The two circular regions are the ascending (right) and descending (left) intersections of the target orbit and \Poincare section.
The green points in~\cref{fig:manifold_poincare} are intersections of the unstable manifold from~\cref{fig:manifold_trajectory} with the \Poincare section \( y = 0 \).
Only a single branch of the invariant manifold intersects with the ascending region of the target orbit.
There are no intersections of the invariant manifold with the descending region of the target orbit.
The numerical values associated with the green points denote the required time of flight along the invariant manifold in non-dimensional units.
The required travel time for a transfer using the unstable invariant manifold is  approximately \( t_f \approx 3.1\) non-dimensional time units.
\begin{figure} 
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{manifold_trajectory} 
                \caption{Position space view of invariant manifold} \label{fig:manifold_trajectory} 
        \end{subfigure}~ %
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{manifold_poincare} 
                \caption{\Poincare intersection of invariant manifold (green), target orbit (blue) and the initial orbit (black)} \label{fig:manifold_poincare} 
        \end{subfigure} 
        \caption{Invariant manifold transfer: An example transfer using the invariant manifolds is shown in both the position and \Poincare spaces.
        The control free transfer from the initial periodic orbit to the target orbit result in a long time of flight. 
    In addition, the manifold only intersects the target orbit on the descending or far side of the moon.}
        \label{fig:invariant_manifold_transfer} 
\end{figure}

@@

% discuss reachability set computation
The optimal trajectories, under the influence of the control input \( \vecbf{u} \), are plotted in red in~\cref{fig:reach_trajectory}.
Initially, the spacecraft is assumed to lie on the periodic orbit.
As a result, the intersection of this periodic orbit with the \Poincare section are two points corresponding to the two crossing of the orbit.
We show the control-free intersection, \( \vecbf{x}_n \), of the periodic orbit on the \Poincare section in~\cref{fig:manifold_poincare,fig:poincare_compare}
The use of the continuous low thrust propulsion expands the reachable set to region bounded by the red markers in~\cref{fig:poincare_compare}.
The reachable set is an ellipsoidal region with a major axis aligned along \( \theta \approx \ang{70} \) as compared to a fixed point without any control input.
\begin{figure} 
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{reach_trajectory} 
                \caption{Position space view of reachability set} \label{fig:reach_trajectory} 
        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{poincare_compare} 
                \caption{\Poincare section view of reachability set} \label{fig:poincare_compare} 
        \end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)

        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{reach_transfer} 
                \caption{Transfer trajectory selected from reachabiltiy set viewed in the position space} \label{fig:reach_transfer} 
        \end{subfigure}~ 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{control_input_l1} 
                \caption{Control input for the selected transfer trajectory} \label{fig:control_l1} 
        \end{subfigure}~
        \caption{\( L_1 \) Reachability set transfer: The low thrust propulsion is used to approximate the reachability set starting from the initial periodic orbit over a fixed time horizon.
        The reachability set is shown in the upper two figures in both the position and \Poincare space.
    From this reachabiilty set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
The optimal control to achieve this transfer is shown in the lower right figure.}
        \label{fig:reachability_set_transfer} 
\end{figure}
